Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-07-05 main_revision.tex
Updates from Overleaf

\subsection{CLM5}
 Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented within this special issue by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019}/. The plant hydraulics model is documented by \cite{kennedy2019}. Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.

\subsection{Updates to biogeochemistry Cycling in the CLM5}
The CLM5 biogeochemistry scheme builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}),  and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{koven2013}, \cite{bonan2012}).

 A full technical description of the CLM5 is beyond the scope of this paper and is available as an appendix to \cite{lawrence2018} and also online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}.

In this paper, we focus on the response to CO$_{2}$ and nitrogen deposition, which are primarily mediated by the biogeochemical processes within the model. The primary changes to the biogeochemical cycling in CLM5 are within the vegetation nitrogen cycle, wherein the model contains three new prognostic elements that are likely to impact responses to fertilization.

Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}). Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). This method, however, appears to be a poor means of capturing variation in photosynthetic capacity in time and space and responses to climatic change (\cite{walker2017}). There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).  LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for a given leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.  $V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5. FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun} and \cite{brzostek2014} and \cite{shi2016}). A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}). Previous versions of CLM model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}). The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation. Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake.   The switching of N uptake to cheaper pathways results in the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}), whereas CLM4 and CLM4.5 predicted N fixation as a linear function of net primary productivity (\cite{cleveland1999}; \cite{wieder2015}).   The FUN model is documented within the CLM technical note at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}.

Thirdly, CLM5 contains the capacity to modify tissue C:N ratios with spatial and temporal variation in the cost of N uptake.  Plant trait databases (e.g. \cite{kattge2011}) typically indicate a wide variability in tissue N concentrations within a single plant functional type (\cite{wright2005}). \cite{ghimire2016} introduced the capacity for variable tissue C:N ratios, whereby the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).  The FUN model used in CLM5 was originally conceived in the context of a static tissue C:N ratio. Carbon is allocated between growth and N uptake in a manner that exactly tracks the target C:N ratio. To allow a variable C:N ratio in the context of FUN, it is necessary to modify how much C is expended on N uptake vs on growth from this default calculation. The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010},\cite{ghimire2016}). Here we include a placeholder algorithm that modifies C:N ratios with 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content. When N is scarce in the environment, less C is spent on uptake, increasing C:N ratios. When plants become N limited and C:N ratios increase, however, expenditure increases. This algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters ($flexcn_{a}$, $flexcn_{b}$ \& $flexcn_{c}$) the sensitivity of which is investigated here.

\subsection{Sites}
We utilized four observational flux tower sites, in order to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, 44.5$^{o}$N -121.6$^{o}$E).  Each site was driven with the available site-specific meteorological data. Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites \cite{bonan2012} for Metolius and Harvard Forest, \cite{fisher2007} for Caxiuan\~a and \cite{bonan2013} for Niwot Ridge.   We note here that our use of a small number of single sites cannot fully describe the environmental conditions under which the model is applied. Instead, our aim is to provide detailed illustrations of how model parameter choices impact a model outputs under a range of realistic forcing conditions. Conducting and analysing these simulations globally is beyond the scope of this study, but investigations of the response of global biogeochemical feedback strength to the important dimensions of parameter space identified by this investigation will provide the basis of future studies.

\subsection{Simulation Protocol}
We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the decomposition of slower cycling carbon and nitrogen pools is increased for the duration of the spin-up, and the pool sizes are modified accordingly at the end of the spin-up - see (\cite{lawrence2018}). We used pre-industrial CO$_{2}$ concentrations (274ppm) and recycled the available site-level half-hourly meteorological inputs for the duration of the simulations. 400 years was sufficient to bring all soil and vegetation carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off. Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and N deposition until 2018.

For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation. Following \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.  Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization. These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in its own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.  We focus here on biogeochemically important inputs. Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around the default state, which (compared to a global parameter sensitivity analysis) allows more straightforward visualization and interpretation of the results.  At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below. We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses. We concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing. 
\subsection{Default parameter determination}
During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets. The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states. We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.

@@

\textbf{Specific Leaf Area}, \emph{`SLA'}, is a parameter common to almost all land surface models, which determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light. In CLM5 high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.

\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass. In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition in the FUN model, respectively. Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.  The carbon allocation in CLM5 uses a fixed-fraction approach, with constant ratios between, leaf, fine roots, stem and coarse root allocation. As such, the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance. Newer codes within the CLM framework (e.g. the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}m \cite{fisher2018vegetation}) account for this issue by removing leaves in negative carbon balance.

 \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the fraction of biomass to stem tissues as a fraction of the leaf allocation. Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass. Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter).

\textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. The FUN model introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake. Therefore, uncertainties associated with the cost functions, in particular, are large. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The six \emph{'n\_costs'} parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.

@@

\textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. As with most land surface schemes, higher leaf N concentrations lead to both higher maintenance respiration costs and higher photosynthetic capacity (in this case via the LUNA model).

\textbf{Slope of the stomatal conductance model}, \emph{medlyn\_slope}. CLM5 replaced the Ball-Berry stomatal conductance scheme with the Medlyn stomatal conductance scheme (\cite{medlyn2011}). The slope parameter determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit, in the absence of plant hydraulics limitations. Thus, is it intrinsically linked to the response of ecosystems to CO$_{2}$.

\textbf{C:N flexibility }, \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c},  as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves. Only the thid parameter (c) had an important impact in this sensitivity analysis, and so we focus here on the impact of that parameter.

@@

\section{Results}
Our analysis used four sites, with eight parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations.  We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen per unit area ($N_{leaf}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

In this section we assess first the impacts of parametric variation on the system state and second the responses to carbon and nitrogen fertilization. We do not focus on the goodness of fit to observations, because our aim is to discern the primary parametric controls over CO$_{2}$ and N fertilization responses, and including extensive data-fitting and calibration experiments (including on parameters mostly unrelated to C and N cycling) would detract from this goal.

\subsection{Overall system state}
The parameter responses of the site-level water and nitrogen limitation (see below) in the control (unfertilized) parameter ensemble are shown in figure \ref{btran state}. GPP, NPP, LAI, $N_{leaf}$ and total vegetation carbon are shown for the Niwot Ridge, (Figure \ref{NR1 state}), Caxiuan\~a (Figure \ref{CAX state}), Harvard Forest (Figure \ref{HVF state}) and Metolius (Figure \ref{MET state}) sites. In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}. Lines of differing colors illustrate responses to variation in the thirteen target parameters, with the central crossover point in each figure corresponding to the default case.

@@

\subsubsection{Parametric control over system state}
In this section, we discuss the impacts of the parameter variation on the initial model state.  Many parameters have both first-order direct impacts, and then further impacts via other parts of the model. Here we attempt to illustrate how these mechanisms operate in CLM5 within the parameter space explored here.

At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on $N_{leaf}$ (Figure \ref{NR1 state}), as expected, since it is used directly in the calculation of both of these quantities. The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf. The impact of SLA on NPP can, however, sometimes be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.

The root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions , 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water (via root length density) and 4) on nutrient limitation through root-mediated N uptake costs. The first two impacts represent the costs of more roots, and the latter two their benefits. CAX and HVF exhibit higher LAI with fewer roots. NWR has large reduction in LAI (and GPP \& NPP) with fewer roots. MET exhibits a mixed response crossing an optimal value. \emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) indicating the impact of mechanism '3' here, for very low root densities.

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon. At NWR, the parameter range crosses a local maximum for vegetation carbon, since increases in NPP are non-linear with respect to leaf investment, resulting in lower vegetation biomass overall. At all other sites, the impact of increasing leaf allocation is to invoke large increases in LAI and NPP. Lower stem allocation and higher LAI increase the N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).

>>>>

\subsection{CLM5}
 Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and ongoing development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented (within this special issue) by \cite{lawrence2018}, a description of the global nitrogen cycling responses to fertilizaion by \cite{wieder2019}. The plant hydraulics model is documented by \cite{kennedy2019}. Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.

\subsection{Updates to biogeochemistry Cycling in the CLM5}
The CLM5 biogeochemistry scheme builds on the initial implementation of carbon-nitrogen interactions in CLM4 (\cite{thornton2007}, \cite{lawrence2011}),  and on modifications to the biogeochemical cycling (notably, a vertically stratified soil decomposition processes, and a slowing of the soil mineral N loss pathways) included in the CLM4.5 (\cite{bonan2012}, \cite{koven2013}). The full technical description of the CLM5 is available as an appendix to \cite{lawrence2018} and also online at \url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}. We refer to this document here, in lieu of repeating large sections of the documentation an as appendix.

In this paper, we focus on the response to CO$_{2}$ and nitrogen deposition, which are primarily mediated by the biogeochemical processes within the model. The primary changes to the biogeochemical cycling in CLM5 are within the vegetation nitrogen cycle. In particular, the new model contains three new prognostic elements that are likely to impact responses to fertilization.

Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}). Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}) but do not capture observed variation in photosynthetic capacity in response to CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).  LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for given leaf nitrogen per unit area ($N_{leaf}$) and environmental boundary conditions, assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.  $V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology.

Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5. FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun}, \cite{brzostek2014} \& \cite{shi2016}). A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}). Previous versions of CLM managed this by reducin the gross photosynthetic flux down to the level at which growth could be supported by N uptake. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}). The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{rastetter2001},  \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation. Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake.   The switching of N uptake to cheaper pathways results in the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}), whereas CLM4 and CLM4.5 predicted N fixation as a linear function of net primary productivity (\cite{cleveland1999}; \cite{wieder2015}).   The FUN model is documented within the CLM technical note at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}.

Thirdly, CLM5 contains the capacity to modify tissue C:N ratios with spatial and temporal variation in the cost of N uptake.  Plant trait databases (e.g. \cite{kattge2011}) typically indicate a wide variability in tissue N concentrations within a single plant functional type (\cite{wright2005}). \cite{ghimire2016} introduced the capacity for variable tissue C:N ratios, whereby the total nitrogen supply in each timestep is partitioned between tissues in proportion to their relative 'demand' terms (ascertained from the target C:N ratio and the carbon allocated to the pool).  The FUN model used in CLM5, however, was originally conceived in the context of a static tissue C:N ratio. Carbon is initially allocated between growth and N uptake in a manner that exactly tracks a target C:N ratio. To allow a variable C:N ratio in the context of FUN, it was necessary to modify how much C is expended on N uptake vs on growth from this default calculation. The degree to which tissue C:N content varies as N becomes limiting is unclear in the literature, and N cycle models typically include semi-heuristic and/or bounded representations of tissue N adjustments under N limited conditions (\cite{zaehle2010}, \cite{ghimire2016}). Here we include a placeholder algorithm that modifies C:N ratios as a function of 1) the environmental cost of N acquisition and 2) the degree to which the plant is already far from its target N content. When N is scarce in the environment, less C is spent on uptake, increasing C:N ratios. When plants become N limited and C:N ratios increase, however, expenditure increases. This algorithm is also documented in the CLM5 technical description ((\url{(https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/index.html)}) and includes three tuning parameters the sensitivity of which is investigated here.

\subsection{Sites}
We utilized four observational flux tower sites, in order to capture a range of environmental conditions, across tropical evergreen broadleaf forest (Caxiuan\~a, CAX -1.9$^{o}$N, -51.4$^{o}$E),  sub-alpine evergreen forest (Niwot Ridge, NWR, 40.3$^{o}$N, -105.4$^{o}$E), temperate deciduous broadleaf forest (Harvard Forest, HVF, 42.5$^{o}$N -72.2$^{o}$E) and seasonally dry evergreen needleleaf forest (Metolius Forest, MET, 44.5$^{o}$N -121.6$^{o}$E).  Each site was driven with the available site-specific meteorological data. Soil texture, color, depth and, plant functional type composition were taken from previous model validation exercises undertaken at these sites \cite{bonan2012} for MET and HVF,  \cite{fisher2007} for CAX and \cite{bonan2013} for NWR.    Conducting and analysing these simulations globally is beyond the scope of this study.  Instead, our aim is to provide detailed illustrations of how model parameter choices impact a model outputs under a range of realistic forcing conditions.

\subsection{Simulation Protocol}
We spun up the default version of the model for 400 years at each site, in 'accelerated spin-up' mode (whereby the decomposition of slower cycling carbon and nitrogen pools is increased for the duration of the spin-up, and the pool sizes are modified accordingly at the end of the spin-up - see (\cite{lawrence2018}). We used pre-industrial CO$_{2}$ concentrations (274ppm) and recycled the available site-level half-hourly meteorological inputs for the duration of the simulations. 400 years was sufficient to bring all soil and vegetation carbon and nitrogen pools into equilibrium. Subsequent to this, parameters were perturbed (see below), and a second spin-up conducted for 180 years for each site, parameter and value combination, with accelerated mode turned off. Restarting from the state at the end of the perturbed parameter spin-ups, we ran 'control' transient simulations for each site and parameter combination starting in 1760 and invoking historical transient CO$_{2}$ and N deposition until 2018. For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation. Following \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.  Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization. These parameters are narrowed down from the much larger complete set of CLM5 parameters using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.  We focus here on biogeochemically important inputs. Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around the default state, which (compared to a global parameter sensitivity analysis) allows more straightforward visualization and interpretation of the results.  At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below. We note that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses, but concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.

\subsection{Default parameter determination}
During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model parameter space using neural-network derived emulators. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the parameter space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystems to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, resulting in marginal areas with many dead grid cells, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed. The parameterization used in the default version of the model consequentially was derived from the mean observed trait values where possible, in combination with targeted manual tuning of parameters (including allocation parameters and nitrogen costs) that were less well constrained by data, where excessively unrealistic values for simulated states were detected. We recognize that this more subjective calibration method is less than ideal. Ongoing studies will focus on more limited-scope calibration attempts to overcome the difficulties described (e.g. \cite{fer2018}).

@@

\textbf{Specific Leaf Area}, \emph{`SLA'} determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies. In CLM5, high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.

\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the ratio of fine root biomass to leaf biomass. In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition, respectively. Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition while higher values imply greater construction and maintenance costs.

 \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the fraction of biomass to stem tissues as a fraction of the leaf allocation. Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass. Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter.

\textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. FUN uses a bulk approach to account for the various costs to plants of low environmental N availability, including carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake. This abstract means that, uncertainties associated with the cost functions, in particular, are large. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The six \emph{'n\_costs'} parameters include, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.

@@

\textbf{Leaf carbon:nitrogen target}, \emph{leafcn}. CLM5 uses the concept of a 'target' leaf C:N ratio, around which flexibility is allowed, given the environmental costs of Nitrogen. As with most land surface schemes, higher leaf N concentrations lead to both higher maintenance respiration costs and higher photosynthetic capacity.

\textbf{Slope of the stomatal conductance model}, \emph{medlyn\_slope}. CLM5 replaced the Ball-Berry stomatal conductance scheme with the Medlyn stomatal conductance scheme (\cite{medlyn2011}). The slope parameter determines the degree of stomatal opening for a given combination of assimilation capacity, CO$_{2}$ and vapour pressure deficit, in the absence of plant hydraulics limitations. It is included here on account of it's critical impact on stomatal responses to CO$_{2}$.

\textbf{C:N flexibility }, the \emph{cn\_flex\_a, cn\_flex\_b, \& cn\_flex\_c}, parameters  as described in the CLM5 technical note, determine the response of tissue C:N ratios to depletion of N in both the environment and the plant tissues themselves. Only the third parameter (c) had an important impact in this sensitivity analysis, and so we focus here on the impact of that parameter.

@@

\section{Results}

\subsection{Overall system state}
In this section we assess first the impacts of parametric variation on the pre-fertilization system state. The parameter responses of the site-level water and nitrogen limitation (see below) in the pre-fertilization parameter ensemble are shown in Fig. \ref{btran state}. GPP, NPP, LAI and $N_{leaf}$ in Fig.\ref{state}. In each of these figures, the y-axis corresponds to the output variable and the x-axis represents the range of parameter perturbation with -1 to +1 corresponding to the range of variation in each parameter, depicted in Table \ref{table_ranges}. Lines of differing colors illustrate responses to variation in the target parameters, with the central crossover point in each figure corresponding to the default case.

@@

\subsubsection{Parametric control over system state}
In this section, we discuss the impacts of the parameter variation on the initial model state.  Many parameters have both first-order direct impacts, and potential indirect effects via modification of the initial state of the system or on other elements of the model. Here we attempt to illustrate how these mechanisms operate in CLM5 within the parameter space explored here.

Specific leaf area (\emph{slatop}) has a large first-order impact on LAI , and on $N_{leaf}$ (Figure \ref{state}), at all sites. \emph{slatop} it is used directly in the calculation of both of these quantities. The impact of SLA on GPP is always positive, despite the thinner leaves having lower N, and thus lower photosynthetic capacity, per unit area of leaf. The impact of SLA on NPP can, however, sometimes be negative (CAX, NWR), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI.

Root:leaf ratio (\emph{froot\_leaf}) has numerous impacts, including 1) on LAI, via tissue allocation fractions, 2) on NPP, via changes in maintenance costs of fine root pools, 3) on the uptake of water, via root length density and 4) on nutrient limitation, via root-mediated N uptake costs. The first two impacts represent the costs of roots, and the latter two their benefits. CAX and HVF exhibit higher LAI with lower root allocation, whereas NWR has large reduction in LAI (and GPP \& NPP) with lower root allocation. MET exhibits a mixed response crossing an optimal value. \emph{froot\_leaf} is a primary control over drought stress at both NWR and MET ( Figure \ref{btran state}) illustrating the impact of low root density on water uptake capacity.

Low values of \emph{stem\_leaf} ratio resulted in higher LAI at all sites, as well as declines in total vegetation carbon (.   The impact of increasing leaf allocation is to invoke large increases in LAI and NPP. Lower stem allocation and higher LAI also increase N demand, reducing $N_{leaf}$ and increase PNU at Niwot Ridge and Metolius in Figure \ref{btran state}).
